---
title: "Quick Start Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Getting Started with ShennongTools

ShennongTools provides a unified interface for running bioinformatics tools through R. This guide will get you up and running in minutes.

## Installation

Install ShennongTools from GitHub:

```{r install}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("zerostwo/shennong-tools")
```

## Basic Workflow

### 1. Load the Package

```{r setup}
library(ShennongTools)

# Initialize the package (one-time setup)
sn_initialize()
```

### 2. Explore Available Tools

```{r explore}
# List all available tools
sn_list_tools()

# Get help for a specific tool
sn_help("samtools")

# See available commands for a tool
sn_show_tool("samtools")
```

### 3. Run Your First Command

The basic syntax is: `sn_run(tool_name, command, ...parameters)`

```{r first_command}
# Example: Convert SAM to BAM
result <- sn_run("samtools", "view",
  input = "alignment.sam",
  output = "alignment.bam",
  flags = "-Sb",
  threads = 4
)

# Check if it was successful
print(result)
```

## Common Patterns

### Quality Control with FastP

```{r fastp}
# Process paired-end FASTQ files
result <- sn_run("fastp", "filter",
  input1 = "sample_R1.fastq.gz",
  input2 = "sample_R2.fastq.gz",
  output1 = "clean_R1.fastq.gz",
  output2 = "clean_R2.fastq.gz",
  html = "fastp_report.html",
  threads = 8
)
```

### RNA-seq Alignment

```{r rnaseq}
# 1. Build HISAT2 index
sn_run("hisat2", "build",
  reference = "genome.fa",
  index_base = "genome_index",
  threads = 8
)

# 2. Align reads
result <- sn_run("hisat2", "align",
  index = "genome_index",
  read1 = "clean_R1.fastq.gz",
  read2 = "clean_R2.fastq.gz",
  bam = "aligned.bam",
  threads = 8
)

# 3. Sort and index BAM
sn_run("samtools", "sort",
  input = "aligned.bam",
  output = "sorted.bam",
  threads = 4
)

sn_run("samtools", "index", input = "sorted.bam")
```

### Peak Calling (ChIP-seq)

```{r chipseq}
# Call peaks with MACS2
result <- sn_run("macs2", "callpeak",
  treatment = "ChIP.bam",
  control = "Input.bam",
  name = "sample",
  format = "BAM",
  gsize = "hs",
  qvalue = 0.05
)
```

## Key Features

### Automatic Installation

Tools are installed automatically on first use:

```{r auto_install}
# First time running a tool - it will be installed automatically
result <- sn_run("seqkit", "stats", input = "sequences.fasta")
```

### Dry Run Mode

Preview commands before running:

```{r dry_run}
# See what command would be executed
result <- sn_run("samtools", "view",
  input = "file.bam",
  output = "filtered.bam",
  flags = "-q 30",
  dry_run = TRUE
)

# Check the rendered command
print(result@rendered_command)
```

### Logging Control

Control output verbosity:

```{r logging}
# Silent mode (minimal output)
result <- sn_run("samtools", "index",
  input = "file.bam",
  log_level = "silent"
)

# Detailed mode (show tool output)
result <- sn_run("samtools", "stats",
  input = "file.bam",
  output = "stats.txt",
  log_level = "normal"
)
```

### Resource Monitoring

Monitor execution resources:

```{r resources}
result <- sn_run("hisat2", "align",
  index = "genome_index",
  read1 = "sample_R1.fastq.gz",
  read2 = "sample_R2.fastq.gz",
  bam = "aligned.bam",
  threads = 8
)

# Check resource usage
cat("Runtime:", sn_get_toolcall_runtime(result), "seconds\n")
cat("Memory:", result@resources$peak_memory_mb, "MB\n")
cat("Success:", sn_is_toolcall_success(result), "\n")
```

## Configuration

### Global Options

Set global preferences:

```{r options}
# Set logging level and log directory
sn_options(
  log_level = "minimal",
  log_dir = "~/analysis_logs"
)

# View current options
sn_options()
```

### Tool Versions

Manage tool versions:

```{r versions}
# Use specific version
result <- sn_run("samtools", "view",
  input = "file.bam",
  version = "1.19.2"
)

# Check installed versions
toolbox <- sn_initialize_toolbox()
tool <- sn_get_tool(toolbox, "samtools")
versions <- sn_get_installed_versions(tool)
print(versions)
```

## Best Practices

1. **Use descriptive file names**: Make your analysis reproducible
2. **Set thread counts**: Optimize for your system's capabilities  
3. **Check outputs**: Always verify that commands completed successfully
4. **Use dry_run**: Preview complex commands before execution
5. **Set log directories**: Keep organized logs for troubleshooting

## Getting Help

- **Package documentation**: `?sn_run`
- **Tool help**: `sn_help("tool_name")`
- **Command help**: `sn_help("tool_name", "command")`
- **Raw tool help**: `sn_help("tool_name", raw = TRUE)`

## Next Steps

- Explore the [Tools Overview](tools-overview.html) for detailed tool documentation
- Learn about [Advanced Usage](advanced-usage.html) features
- Check out [YAML Specification](yaml-specification.html) for custom tools

Happy analyzing! 🧬 
